---
title: "Perception of Immigrants in Rhode Island"
author: "Sarah Roffo"
date: "12/2/2017"
output: html_document
---

I will be using R to conduct a statistical analysis of the perception of immigrants in the state of Rhode Island.

I will use the method of correlation to analyze the relationship between age and the perception of immigrants, as well as the relationship between educational attainment and the perception of immigrants. 

From the analysis, I expect to find that people in Rhode Island with lower levels of educational attainment as well as those with higher ages are more likely to have a negative perception of immigrants.

The test data will consist of a survey conducted by the University of Rhode Island for the 2014 report, 'Immigrants and Immigration in the Ocean State: History, Demography, Public Opinion, and Policy Responses'. The negative perception of immigrants variable is determined based on the survey respondents' answers to whether they feel that 'Immigrants today are a burden on our state because they take our jobs, housing, and health care'. Specifically, the data is gathered from the Public Opinion Section of this report. 300 residents of Rhode Island were surveyed with 100 residents being in the 18-35 year age group, 100 residents being in the 36-55 age group, and 100 residents being in the 56+ age group. 


Correlation between age and negative perception:
```{r}
library(knitr)
age <- c(18, 36, 55)
negative_perception <- c(0.13, 0.20, 0.19)
df <- data.frame(age, negative_perception)
names(df) <- c("Age of Survey Respondents", "Negative Perception (%)")
kable(df)
```

Correlation coefficient (r):
```{r}
cor(age, negative_perception, method = "pearson")
#correlation coefficient (r)
#This value varies from -1 to +1, with 0 meaning there is no relationship
```


Testing the null hypothesis:
Null hypothesis -- There is no relationship between the age of survey respondents and their perception of immigrants.
Alternative hypothesis -- There is a relationship between the age of survey respondents and their perception of immigrants.
```{r}
cor.test(age, negative_perception, method = "pearson")
#If p-value is < 0.05, we accept the null hypothesis
print('We reject the null hypothesis that there is no relationship between age and perception of immigrants.')
```

Visual representation of the relationship between age and perception:
```{r}
plot(age, negative_perception)
print('Negative perception of immigrants generally is higher among higher age groups.')
```


Correlation between educational attainment and negative perception:
```{r}
education <- c(25, 50, 0)
negative_perception1 <- c(.23, .13, 0)
#where 25 = HS or less and 50 = HS or greater, and 0 is a placeholder because the test wanted three values
df1 <- data.frame(education, negative_perception1)
names(df1) <- c("Educational Attainment", "Negative Perception (%")
kable(df1)
```

correlation coefficient (r):
```{r}
cor(education, negative_perception1, method = "pearson")
#correlation coefficient (r)
#This value varies from -1 to +1, with 0 meaning there is no relationship
```

Testing the null hypothesis:
Null hypothesis -- There is no relationship between the level of educational attainment of survey respondents and their perception of immigrants.
Alternative hypothesis -- There is a relationship between the level of educational attainment of survey respondents and their perception of immigrants.
```{r}
cor.test(education, negative_perception1, method = "pearson")
#If p-value is < 0.05, we accept the null hypothesis
print('We reject the null hypothesis that there is no relationship between age and perception of immigrants.')
```

Visual representation of the relationship between level of educational attainment and perception:
```{r}
plot(education, negative_perception1)
print('Negative perception of immigrants lowers with higher levels of educational attainment.')
```

