---
title: "Lab 6"
author: "Sarah Roffo"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

Complete the following exercises in Chapter 6 (Analysis of Variance) of the textbook pages 199-203. For each question, you need to specify the null hypothesis and why you accept or reject the null hypothesis.


## Question 1
This is Exercise 2 in Chapter 6 of the Textbook [R].

### Problem
Assume that an analysis of variance is conducted for a study where there are $N = 50$ observations and $k = 5$ categories. Fill in the blanks in the following ANOVA table:

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
| Between  |                                      |      116.3     |                 |
| Within   |       2000      |                    |                |                 |
| Total    |                 |                    |                |                 |



### Solution
```{r}
N <- 50
k <- 5
#F = (BSS/(k - 1))/(WSS/(n - k))

## define the variables here, and use the imbedded r code like `r variable` to fill in the blanks.  

qf(p = 0.95, df1 = 4, df2 = 45)

#Null Hypothesis: The means are the same
#Alternative Hypothesis: The means are different

print('We reject the null hypothesis because the means are different')
```

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
| Between  |`r 116.3*(k - 1)`|       `r k-1`      |      116.3     |  `r (116.3/(k - 1)/(2000/(N - k)))`|
| Within   |       2000      |       `r N-k`      |`r 2000/(N - k)`|                 |
| Total    |`r 116.3*(k - 1) + 2000`|                    |              |                 |




## Question 2
This is Exercise 6 in Chapter 6 of the Textbook [R].

### Problem 
Is there a significant difference between the distances moved by low- and high-income individuals? Twelve respondents in each of the income categories are interviewed, with the following results for the distances associated with residential moves:

| Respondent  | Low income  | High income |
|-------------|-------------|-------------|
|     1       |      5      |     25      |
|     2       |      7      |     24      |
|     3       |      9      |     8       |
|     4       |      11     |     2       |
|     5       |      13     |     11      |
|     6       |      8      |     10      |
|     7       |      10     |     10      |
|     8       |      34     |     66      |
|     9       |      17     |     113     |
|     10      |      50     |     1       |
|     11      |      17     |     3       |
|     12      |      25     |     5       |
|     Mean    |      17.17  |     23.17   |
|  Std. dev.  |      13.25  |     33.45   |


Test the null hypothesis of homogeneity of variances by forming the ratio $s_1^2 / s_2^2$ which has an F-ratio with $n_1 – 1$ and $n_2 – 1$ degrees of freedom. Then use ANOVA (with $\alpha = 0.10$) to test whether there are differences in the two population means. Set up the null and alternative hypotheses, choose a value of α and a test statistic, and test the null hypothesis. What assumption of the test is likely not satisfied?

### Solution
```{r}
L_income <- c(5, 7, 9, 11, 13, 8, 10, 34, 17, 50, 17, 25)
H_income <- c(25, 24, 8, 2, 11, 10, 10, 66, 113, 1, 3, 5)
Distances <- c(L_income, H_income)
Group <- c(rep("A", times = 12), rep("B", times = 12))
df <- data.frame(Distances, Group)
m <- aov(Distances~Group, data = df)
summary(m)

qf(0.95, df1 = 1, df2 = 22)

#Null Hypothesis: There is no significant difference in the distances moved by low income and high income individuals
#Alternative Hypothesis: There is a significant difference in the distanced moved by low income and high income individuals

#Accept Null if Pr(>F) > 0.05

print('We accept the null hypothesis because there is no significant difference in the distances moved by low income and high income individuals')
```


## Question 3
This is Exercise 9 in Chapter 6 of the Textbook [R].

### Problem
A sample is taken of incomes in three neighborhoods, yielding the following data: 

|          |        A        |          B         |       C        | Overall (Combined sample) |
|----------|-----------------|--------------------|----------------|---------------------------|
| N        |        12       |          10        |        8       |             30            |
| Mean     |       43.2      |          34.3      |        27.2    |             35.97         |
| Std. dev.|       36.2      |          20.3      |        21.4    |             29.2          |


Use analysis of variance (with α = 0.05) to test the null hypothesis that the means are equal.


### Solution
```{r}
library(MASS)
A <- mvrnorm(n = 12, mu = 43.2, Sigma = (36.2 * 36.2), empirical = TRUE)
B <- mvrnorm(n = 10, mu = 34.3, Sigma = (20.3 * 20.3), empirical = TRUE)
C <- mvrnorm(n = 8, mu = 27.2, Sigma = (21.4 * 21.4), empirical = TRUE)
incomes <- c(A, B, C)
Group <- c(rep("A", times = 12), rep("B", times = 10), rep("C", times = 8))
df <- data.frame(incomes, Group)
m <- aov(incomes~Group, data = df)
summary(m)

#Null Hypothesis: The means are equal
#Alternative Hypothesis: The means are not equal

#Reject Null if Pr(>F) < 0.05

print('We accept the null hypothesis because all of the means are equal')
```



## Question 4
This is Exercise 10 in Chapter 6 of the Textbook [R].

### Problem
Use the Kruskal–Wallis test (with α = 0.05) to determine whether you should reject the null hypothesis that the means of the four columns of data are equal:

|   Col 1  |       Col 2     |        Col 3       |       Col 4    |
|----------|-----------------|--------------------|----------------|
|   23.1   |       43.1      |        56.5        |       10002.3  |
|   13.3   |       10.2      |        32.1        |       54.4     |
|   15.6   |       16.2      |        43.3        |       8.7      |
|   1.2    |       0.2       |        24.4        |       54.4     |


### Solution
```{r}
A <- c(23.1, 13.3, 15.6, 1.2)
B <- c(43.1, 10.2, 16.2, 0.2)
C <- c(56.5, 32.1, 43.3, 24.4)
D <- c(10002.3, 54.4, 8.7, 54.4) 

columns <- c(A, B, C, D)
Group <- c(rep("A", times = 4), rep("B", times = 4), rep("C", times = 4), rep("D", times = 4))
df <- data.frame(columns, Group)
kruskal.test(columns~Group, data = df)

#If p-value greater than .05 we cannot reject the null Hypothesis

#Null Hypothesis: The means of the four columns of data are equal
#Alternative Hypothesis: The means of the four columns of data are not equal

print('We accept the null hypothesis')
```


## Question 5
This is Exercise 12 in Chapter 6 of the Textbook [R].

### Problem
A researcher wishes to know whether distance traveled to work varies by income. Eleven individuals in each of three income groups are surveyed. The resulting data are as follows (in commuting miles, one-way):

```{r}
## This is the script to generate the table. Do not write your answer inside in this block.
Observations <- seq(1:11)
Low <- c(5,4,1,2,3,10,6,6,4,12,11)
Medium <- c(10,10,8,6,5,3,16,20,7,3,2)
High <- c(8,11,15,19,21,7,7,4,3,17,18)

df <- data.frame(Observations,Low,Medium,High)
library(knitr)
kable(df)
```

Use analysis of variance (with α = 0.05) to test the hypothesis that commuting distances do not vary by income. Also evaluate (using R and the Levene test) the assumption of homoscedasticity. Finally, lump all of the data together and produce a histogram, and comment on whether the assumption of normality appears to be satisfied.


### Solution
```{r}
library(reshape2)
library(car)

A <- c(5,4,1,2,3,10,6,6,4,12,11)
B <- c(10,10,8,6,5,3,16,20,7,3,2)
C <- c(8,11,15,19,21,7,7,4,3,17,18)

Income <- c(A, B, C)
Group <- c(rep("A", times = 11), rep("B", times = 11), rep("C", times = 11))
df <- data.frame(Income, Group)
m <- aov(Income~Group, data = df)
summary(m)
#Null Hypothesis: Commuting Distances do not vary by income
#Alternative Hypothesis: Commuting Distances do vary by income
#Reject Null if Pr(>F) < 0.05

print('We reject the Null Hypothesis')

##Assumption of homoscedasticity
dataset <- melt(data = df)
leveneTest(Income~Group, dataset)

print('We accept the Null Hypothesis')

hist(df$Income)
shapiro.test(df$Income)

#p < 0.05 indicates that the population is likely not normally distributed
print('The population is likely not normally distributed')
```


## Question 6
This is Exercise 13 in Chapter 6 of the Textbook [R].

### Problem
Data are collected on automobile ownership by surveying residents in central cities, suburbs and rural areas. The results are:

|                      | Central cities  |      Suburbs       |  Rural areas   |
|----------------------|-----------------|--------------------|----------------|
|Number of observations|      10         |        15          |       15       |
|      mean            |      1.5        |        2.6         |       1.2      |
|      Std. dev        |      1.0        |        1.1         |       1.2      |  
|Overall mean: 1.725   |                 |                    |                |  
|Overall std.dev: 1.2  |                 |                    |                |   


Test the null hypothesis that the means are equal in all three areas.

### Solution
```{r}
library(MASS)
A <- mvrnorm(n = 10, mu = 1.5, Sigma = (1 * 1), empirical = TRUE)
B <- mvrnorm(n = 15, mu = 2.6, Sigma = (1.1 * 1.1), empirical = TRUE)
C <- mvrnorm(n = 15, mu = 1.2, Sigma = (1.2 * 1.2), empirical = TRUE)
place <- c(A, B, C)
Group <- c(rep("A", times = 10), rep("B", times = 15), rep("C", times = 15))
df <- data.frame(place, Group)
m <- aov(place~Group, data = df)
summary(m)

#Null Hypothesis: Means are equal in all three areas
#Alternative Hypothesis: Means are not equal in all three areas

#Reject Null if Pr(>F) < 0.05
print('We reject the null hypothesis, means are not equal')
```

