---
title: "Lab 9"
author: "Sarah Roffo"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

Complete the following exercises in Chapter 9 (More on Regression) of the textbook pages 286-287. 


## Question 1

This is Exercise 7 in Chapter 9 of the Textbook [R].

The following results were obtained from a regression of $n = 14$ housing prices (in dollars) on median family income, size of house, and size of lot:

```{r, echo=FALSE}
library(knitr)
F <- (4234/3)/(3487/(14 - 4))
col1 <- c("Regression SS:","Residual SS:","Total SS:")
col2 <- c("4234", "3487", 4234 + 3487)
col3 <- c("3", 14 - 4, 14 - 1)
col4 <- c(4234/3, 3487/10, "")
col5 <- c(F, "", "")
df <- cbind(col1,col2,col3,col4,col5)
colnames(df) <- c("","Sum of squares","df","Mean square","F")
kable(df)
```

```{r, echo=FALSE}
library(knitr)
col1 <- c("Median family income","Size of house (sq.ft)","Size of lot (sq.ft)","Constant")
col2 <- c("1.57","23.4","-9.5","40,000")
col3 <- c("0.34","11.2","7.1","1000")
col4 <- c("1.3","2.9","11.3","")
df <- cbind(col1,col2,col3,col4)
colnames(df) <- c("","Coefficient (b)","Standard error (sb)","VIF")
kable(df)
```


### Question 1(a)
Fill in the blanks.

```{r}

```


### Question 1(b)
What is the value of $r^2$? 

```{r}
4234/(4234 + 3487)
```


### Question 1(c)
What is the standard error of the estimate? 

```{r}
sqrt(3487/10)
```


### Question 1(d)
Test the null hypothesis that $R^2 = 0$ by comparing the $F-statistic$ from the table with its critical value. 
```{r}
crit <- qf(0.95, 3, 10)
crit
print('We reject the null hypothesis because the critical value of 3.410534 is less than the F-statistic value of 4.04741420514291')
```


### Question 1(e)
Are the coefficients in the direction you would hypothesize? If not, which coefficients are opposite in sign from what you would expect? 

```{r}
print('No, I would expect the coefficients to all increase with higher median family income. The coefficient that is opposite in sign from what I would expect is the size of lot (sq. ft) coefficient.')
```


### Question 1(f)
Find the $t-statistics$ associated with each coefficient, and test the null hypotheses that the coefficients are equal to zero. Use $ \alpha = 0.05$, and be sure to give the critical value of $t$. 

```{r}
income_t <- 1.57/0.34
house_t <- 23.4/11.2
lot_t <- (-9.5)/7.1

critical_t <- qt(0.975,10)
income_t > critical_t
print('We reject the null -- median family income coefficient is not equal to zero.')
house_t > critical_t
print('We fail to reject the null hypothesis for the size of house coefficient.')
lot_t > critical_t
print('We fail to reject the null hypothesis for the size of lot coefficient.')
```


### Question 1(g)
What do you conclude from the variance inflation factors (VIFs)? What (if any) modifications would you recommend in light of the VIFs? 

```{r}
print('From the variance inflation factors, I conclude that the size of lot factor varies the most. I would recommend removing this factor if any are to be removed.')
```


### Question 1(h)
What is the predicted sales price of a house that is 1500 square feet, on a lot 60´×100´, and in a neighborhood where the median family income is $40,000?

```{r}
predicted_sales_price <- (1500 * 23.4) - (6000 * 9.5) + (40000 * 1.57)
predicted_sales_price
print('The predicted sales price of this house is $40,900.')
```



## Question 2

This is Exercise 10 in Chapter 9 of the Textbook [R].

### Question 2(a)
Using R and the [Hypothetical UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv), construct a regression equation using housing price as the dependent variable, and bedrooms, bathrooms, date built, garage, fireplace, floor area, and whether the home is detached as the independent variables. Investigate the importance of multicollinearity and outliers. Comment on the weaknesses of this specification, and on the results. 

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
UK_Housing <- read.csv(url)

price <- UK_Housing$price
garage <- UK_Housing$garage
bedrooms <- UK_Housing$bedrooms
bathrooms <- UK_Housing$bathrooms
date_built <- UK_Housing$datebuilt
floor_area <- UK_Housing$floorarea
detached <- UK_Housing$detached
fireplace <- UK_Housing$fireplace
m <- lm(price ~ garage + bedrooms + bathrooms + date_built + floor_area + detached + fireplace)
summary(m)
plot(m)
print('There appears to be an outlier coefficient that varies from the norm on the regression line at the upper end of the line.')
```


### Question 2(b)
Attempt to improve the regression equation found in (a). Justify your decisions in constructing and carrying out the analysis.

```{r}
improved_regression <- step(m, direction = "backward")
summary(improved_regression)
print('I used the backward stepwise regression because it will remove the underperforming independent variables, if there are any. The improved model excluded the bedrooms independent variable.')
```



## Question 3

This is Exercise 11 in Chapter 9 of the Textbook [R].

### Question 3(a)

Using R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) described in Section 1.9.2, construct a regression equation using housing sales price as the dependent variable, and number of bedrooms, lot size, finished square footage in the house, age of house, and number of bathrooms, as the independent variables. Investigate the importance of multicollinearity and outliers. Comment on the weaknesses of this specification, and on the results. 

```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
Milwaukee <- read.csv(url)
price <- Milwaukee$SalePrice
bedrooms <- Milwaukee$Bedrms
lot_size <- Milwaukee$LotSize
square_footage <- Milwaukee$FinSqft
age <- Milwaukee$Age
bathrooms <- Milwaukee$Baths
m <- lm(price ~ bedrooms + lot_size + square_footage + age + bathrooms)
summary(m)
plot(m)
print('There dont appear to be any outlier coefficients that vary severely from the norm on the regression line.')
```


### Question 3(b)
Attempt to improve the regression equation found in (a). Justify your decisions in constructing and carrying out the analysis.

```{r}
improved_regression <- step(m, direction = "backward")
summary(improved_regression)
print('I used the backward stepwise regression because it will remove the underperforming independent variables, if there are any. No independent variables were excluded in this attempt to improve the equation.')
```

