---
title: "Lab 8"
author: "Sarah Roffo"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
N <- 24
x <- 2.3 - 1.7
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|      `r x`      |          1         |     `r x/1`    |`r x/(1.7/(24 - 2))`|
| Residual |       1.7       |    `r N - 2`       |`r 1.7/(24 - 2)`|                 |
| Total    |       2.3       |    `r N - 1`       |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
trips <- ((0.19 * 50) + 0.46)
trips
print('The predicted number of weekly shopping trips for someone making $50,000/year is 9.96 trips.')
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?

```{r}
print('Every additional $1,000 in annual income increases the chances of a weekly shopping trip by 0.19.')
```


### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
f <- x/(1.7/(24 - 2))
qf(0.95, 1, 22)
print('Yes, the regression coefficient is significantly different from 0. I know this from comparing the F value to f critical.')
```



### Question 1(e)
What is the value of the correlation coefficient?
```{r}
correlation_coefficient <- sqrt((2.3 - 1.7)/2.3)
correlation_coefficient
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
snowfall <- c(36, 78, 11, 45)
elevation <- c(400, 800, 200, 675)
df <- data.frame(snowfall, elevation)
m <- lm(snowfall ~ elevation, data = df)
summary(m)
m$coefficients
#intercept = -7.63720599
#slope coefficient = 0.09665004
```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
summary(m)
#Residual standard error = 11.25
```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
conf <- qf(0.95, 1, 2)
conf
print('We accept the null hypothesis that the regression coefficient associated with the independent variables is equal to zero.')
```


### Question 2(d)
Find the value of $r^2$.
```{r}
(cor(snowfall, elevation))^2
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
library(knitr)
df$Predicted_Values <- m$fitted.values
df$Residuals <- m$residuals
kable(df)
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
anova(m)
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|     2047.77     |         1          |     2047.77    |     16.173      |
| Residual |     253.23      |         2          |     126.61     |                 |
| Total    |`r 2047.77 + 253.23`|      3          |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
df <- data.frame(snowfall, elevation)
plot(snowfall, elevation)
m <- lm(snowfall ~ elevation, data= df)
abline(m, col = "blue")
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
Milwaukee_Sales <- read.csv(url)
price <- Milwaukee_Sales$SalePrice
lot <- Milwaukee_Sales$LotSize
df <- data.frame(price, lot)
m <- lm(price ~ lot, data = df)
summary(m)
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
UK_Housing <- read.csv(url)
price <- UK_Housing$price
bedrooms <- UK_Housing$bedrooms
df <- data.frame(price, bedrooms)
m <- lm(price ~ bedrooms, data = df)
summary(m)
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
price <- UK_Housing$price
bathrooms <- UK_Housing$bathrooms
df <- data.frame(price, bathrooms)
m <- lm(price ~ bathrooms, data = df)
summary(m)

print('The results show that the price of a home is related to the number of bathrooms in it.')
```


## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
library(leaflet)
url = "http://spatial.binghamton.edu/geog533/data/hometown.csv"
hometown <- read.csv(url)
m <- leaflet() 
m <- addTiles()
m <- addMarkers(lng = 43.232059, lat = -76.30052, popup = "Sarah Roffo")
nrow(hometown)
my_info <- c("Roffo", "Sarah", "grads", "43.232059", "-76.300552", "NA", "Sarah Roffo", "NA")
web_map <- as.data.frame(rbind(as.matrix(hometown), my_info))

```


