---
title: "Lab 10"
author: "Sarah Roffo"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r}
library(UScensus2010)
if(!require(UScensus2010county)) install.county("osx")
library(UScensus2010county)
data(new_york.county10)
shp <- new_york.county10
plot(shp)
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r}
library(tmap)
qtm(shp)
```


### Question 1(c)	
How many counties in New York State?
```{r}
df1 <- shp@data
nrow(df1)
print('There are 62 counties in New York State')
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
countyfips[countyfips$countyname == "broome county",]
print('The fips code of Broome County is 007 -- the last three digits of the fips column')
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
df1 <- shp@data
min(df1$P0010001)
max(df1$P0010001)
mean(df1$P0010001)
median(df1$P0010001)
library(e1071)
skewness(df1$P0010001)
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(df1$P0010001)
boxplot(df$P0010001)
```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)


### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
library(UScensus2010)
if(!require(UScensus2010tract)) install.tract("osx")
library(UScensus2010tract)
data(new_york.tract10)
shp <- new_york.tract10
plot(shp)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
df2 <- shp@data
pop <- df$P0010001
sum(pop)
print('The total population of New York State based on census tracts is 19,378,102.')
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
df3 <- df2[df2$county == "007",]
shp <- df3
plot(df3)
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
df3 <- df2[df2$county == "007",]
total.pop <- sum(df3$P0010001)
sum(total.pop)
print('The total population of Broome County is 200,600.')
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(df3$P0010001)
boxplot(df3$P0010001)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}
library(raster)
library(ncdf4)
?head
```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)
```{r}

```

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
NDVI <- nc_open("NDVI.nc")
NDVI
NDVI_RB <- brick("NDVI.nc")
NDVI_RB
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(NDVI_RB)
ncol(NDVI_RB)
ncell(NDVI_RB)
#bands?
res(NDVI_RB)
extent(NDVI_RB)
bbox(NDVI_RB)
projection(NDVI_RB)
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}

```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}

```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r}

```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}

```


